### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@7c7d7768c1e549485c1655de1f1c4b42888ecbdb 🚀

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune
strategies)\textsuperscript{\textbf{Banerjee2020NovIns?}} creates a
landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian
continents,\textsuperscript{\textbf{Shi2014DeeDiv?}} as one such
hotspot; interestingly, a recent
study\textsuperscript{\textbf{Kettenburg2022FulGen?}} reported a novel
and highly divergent lineage of nobecoviruses from Madagascar-endemic
pteropid bat species (\emph{Pteropus rufus} and \emph{Rousettus
madagascariensis}), again supporting the predictive power of the
coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become
epidemics)\textsuperscript{\textbf{Plowright2017PatZoo?}}. It must be
noted that the assesment of risk based on the GMTC mechanisms does not
account for human presence; for this reason, it represents ``potential''
level of risk, which must be re-evaluated in the light of human
presence. As a proxy for both, we finally overlaid the risk component
from the composite map (see above) with the proportion of built land, as
a proxy for a mix of habitat disturbance, potential for bat synanthropy
or contact with bridge hosts like
livestock,\textsuperscript{\textbf{Rulli2021LanCha?},\textbf{Cui2019OriEvo?}}
and human population density and
connectivity\textsuperscript{\textbf{Plowright2017PatZoo?},\textbf{Muylaert2022PreFut?},\textbf{Hassell2017UrbDis?}}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
Similarly, the SARS-CoV outbreak began in Guangdong province in 2002,
reaching humans through small carnivore bridge hosts, but was eventually
traced back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan
province;\textsuperscript{\textbf{Hu2017DisRic?}} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{\textbf{Wang2018SerEvi?}} MERS-CoV was first
detected in Jordan, but is widespread in camels in East Africa and the
Middle East, and may have reached its bridge host decades earlier than
originally supposed;\textsuperscript{\textbf{Muller2014MerCor?}} as a
result, the geography of the original bat-to-camel transmission is still
widely regarded as uncertain. All of these are broadly consistent with
the risk factors we identify. Notably, India and west Africa are
additional hotspots that have yet to experience the emergence of a bat
coronavirus into human populations, but may still be at
risk---particularly given known gaps in bat
surveillance,\textsuperscript{\textbf{Cohen2022SamStr?}} and a dense
population in both regions with global connectivity. In any of these
regions, surveillance on viral reservoirs can be paired with targeted
monitoring of high-risk human populations (i.e., those with regular
wildlife contact)\textsuperscript{\textbf{Xu2004EpiClu?}} for maximum
impact.

@@

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{\textbf{Peixoto2018SynEco?},\textbf{Simmons2020BatSpe?}}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to
humans.\textsuperscript{\textbf{Kasso2013EcoEco?}} Over two-thirds of
bats are known to be either obligate or facultative insectivores,
therefore actively contributing for agricultural pest
control,\textsuperscript{\textbf{Voigt2016BatAnt?},\textbf{Williams-Guillen2008BatLim?}}
and vectors of pathogens that put a risk on human
health;\textsuperscript{\textbf{Gonsalves2013MosCon?},\textbf{Gonsalves2013MosInf?}}
some other species are essential links in many seed-dispersal
networks.\textsuperscript{\textbf{Mello2011MisPar?}} However, many of
these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{\textbf{Olival2017HosVir?}} bat
coronaviruses,\textsuperscript{\textbf{Anthony2017GloPat?}} and even
specifically
betacoronaviruses\textsuperscript{\textbf{Becker2022OptPre?}} in both
the Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin
America.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Caraballo2022CroTra?}}
In contrast, in southeast Asia, host richness and viral distinctiveness
are high but sharing is low; this suggests a different type of
evolutionary dynamics that could generate high local diversity of
viruses through host switching and viral recombination (see
\emph{e.g.},\textsuperscript{\textbf{Latinne2020OriCro?}} as well as the
discovery of recombinant viruses with genetic material from both the
SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{\textbf{Wu2021ComSur?}}Both of these regions
are priority areas for sampling, especially given predictions that they
contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Cohen2022SamStr?}}
However, both the evolutionary and ecological aspects of emergence risk
are higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of elevated cross-species transmission
unique to the
region.\textsuperscript{\textbf{Carlson2022CliCha?},\textbf{Muylaert2022PreFut?}}

Our trivariate additive mapping of components of risk
(fig.~\ref{fig:trivariate}) aims to elicit the complexity of spatial
cross-species transmission risk beyond the mere presence or absence of
the pathogen host in a specific location. By considering coevolutionary
factors such as viral sharing and host uniqueness, we suggest insights
that can aid in identifying potential locations for surveillance of
betacoronavirus circulation and assessing the risk of cross-species
transmission to other mammals. In communities characterized by diverse
but unique host populations, with limited viral sharing between them, we
could encounter viruses that specialize in targeting the immune system
of specific hosts. This implies a low likelihood of infecting novel
hosts but, once locally introduced into a new host (either a new
species, or an immunologically naïve population), the specialized virus
could spread relatively easily due to encountering little immune
resistance (CITE PLOWRIGHT). With the right combination of viral traits,
such as low disease-induced mortality or high transmission rate, this
could lead to successfully spread within the new host community.
However, while high adaptation to a specific host can be advantageous,
it may also lead to maladaptation when the pathogen encounters a new
unsuitable host, potentially resulting in its extinction.

\begin{quote}
Plowright RK, Foley P, Field HE, Dobson AP, Foley JE, Eby P, Daszak P.
Urban habituation, ecological connectivity and epidemic dampening: the
emergence of Hendra virus from flying foxes (Pteropus spp.). Proceedings
of the Royal Society B: Biological Sciences. 2011 Dec
22;278(1725):3703-12.
\end{quote}

Bats---and the spillover of their viruses---are also sensitive to
anthropogenic factors others than climate change, including
deforestation and other kinds of habitat loss, increased stress, and
greater contact with potential bridge hosts like domesticated
species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
This represents a challenge for both conservation strategies and
pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
identifying areas at risk, and protecting the health of bats and
ecosystems within those zones, can be a win-win intervention for
both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
As we scale these predictions down in space to finer spatial resolutions
to guide public health actions,\textsuperscript{27} the incorporation of
human activity predictors will become more
importyant.\textsuperscript{28}

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{29} especially for
beta-coronaviruses.\textsuperscript{13} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

>>>>

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{27} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian continents,\textsuperscript{28} as one
such hotspot; interestingly, a recent study\textsuperscript{29} reported
a novel and highly divergent lineage of nobecoviruses from
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
\emph{Rousettus madagascariensis}), again supporting the predictive
power of the coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become epidemics)\textsuperscript{1}. It
must be noted that the assesment of risk based on the GMTC mechanisms
does not account for human presence; for this reason, it represents
``potential'' level of risk, which must be re-evaluated in the light of
human presence. As a proxy for both, we finally overlaid the risk
component from the composite map (see above) with the proportion of
built land, as a proxy for a mix of habitat disturbance, potential for
bat synanthropy or contact with bridge hosts like
livestock,\textsuperscript{30,31} and human population density and
connectivity\textsuperscript{1,32,33} (fig.~\ref{fig:compound}).
Accounting for these factors, most of South America and Europe are at
comparatively lower risk, as--although densely populated--settlements
tend to be in areas with lower potential risk. Conversely, regions like
Malaysia and the northern coast of Australia have a high evolutionary
risk component, but should represent a relatively lower effective risk
due to low human density. However, southeast Asia, the Indian
subcontinent, and scattered hotspots in sub-Saharan Africa are at high
risk due to the overlap between human populations and natural
opportunities for cross-species transmission of betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{34} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
first detected in Jordan, but is widespread in camels in East Africa and
the Middle East, and may have reached its bridge host decades earlier
than originally supposed;\textsuperscript{36} as a result, the geography
of the original bat-to-camel transmission is still widely regarded as
uncertain. All of these are broadly consistent with the risk factors we
identify. Notably, India and west Africa are additional hotspots that
have yet to experience the emergence of a bat coronavirus into human
populations, but may still be at risk---particularly given known gaps in
bat surveillance,\textsuperscript{24} and a dense population in both
regions with global connectivity. In any of these regions, surveillance
on viral reservoirs can be paired with targeted monitoring of high-risk
human populations (i.e., those with regular wildlife
contact)\textsuperscript{37} for maximum impact.

@@

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{\textbf{Peixoto2018SynEco?},\textbf{Simmons2020BatSpe?}}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to humans.\textsuperscript{38} Over
two-thirds of bats are known to be either obligate or facultative
insectivores, therefore actively contributing for agricultural pest
control,\textsuperscript{39,40} and vectors of pathogens that put a risk
on human health;\textsuperscript{41,42} some other species are essential
links in many seed-dispersal networks.\textsuperscript{43} However, many
of these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{19} bat coronaviruses,\textsuperscript{11} and
even specifically betacoronaviruses\textsuperscript{20} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{11,44} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{45}Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{20,24} However,
both the evolutionary and ecological aspects of emergence risk are
higher in southeast Asia---a fact that will only become more relevant,
as bats track shifting climates and exchange viruses with other species,
creating a hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{32,46}

Our trivariate additive mapping of components of risk
(fig.~\ref{fig:trivariate}) aims to elicit the complexity of spatial
cross-species transmission risk beyond the mere presence or absence of
the pathogen host in a specific location. By considering coevolutionary
factors such as viral sharing and host uniqueness, we suggest insights
that can aid in identifying potential locations for surveillance of
betacoronavirus circulation and assessing the risk of cross-species
transmission to other mammals. In communities characterized by diverse
but unique host populations, with limited viral sharing between them, we
could encounter viruses that specialize in targeting the immune system
of specific hosts. This implies a low likelihood of infecting novel
hosts but, once locally introduced into a new host (either a new
species, or an immunologically naïve population), the specialized virus
could spread relatively easily due to encountering little immune
resistance.\textsuperscript{\textbf{Plowright2011Urban?}} With the right
combination of viral traits, such as low disease-induced mortality or
high transmission rate, this could lead to successfully spread within
the new host community. However, while high adaptation to a specific
host can be advantageous, it may also lead to maladaptation when the
pathogen encounters a new unsuitable host, potentially resulting in its
extinction.

Bats---and the spillover of their viruses---are also sensitive to
anthropogenic factors others than climate change, including
deforestation and other kinds of habitat loss, increased stress, and
greater contact with potential bridge hosts like domesticated
species.\textsuperscript{30,47--49} This represents a challenge for both
conservation strategies and pandemic prevention,\textsuperscript{50} but
identifying areas at risk, and protecting the health of bats and
ecosystems within those zones, can be a win-win intervention for
both.\textsuperscript{51,\textbf{Hopkins2021HowIde?},\textbf{Plowright2021Lan?}}
As we scale these predictions down in space to finer spatial resolutions
to guide public health actions,\textsuperscript{32} the incorporation of
human activity predictors will become more
importyant.\textsuperscript{52}

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{53} especially for
beta-coronaviruses.\textsuperscript{13} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

\leavevmode\hypertarget{ref-Banerjee2020Novel}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Shi2014Deep}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022Full}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Rulli2021Landuse}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

\leavevmode\hypertarget{ref-Cui2019Origin}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
\textbf{17}, 181--192 (2019).}

\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Hassell2017Urbanization}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
Urbanization and disease emergence: Dynamics at the
wildlife--livestock--human interface. \emph{Trends in ecology \&
evolution} \textbf{32}, 55--67 (2017).}

\leavevmode\hypertarget{ref-Hu2017Discovery}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
bat SARS-related coronaviruses provides new insights into the origin of
SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}

\leavevmode\hypertarget{ref-Wang2018Serological}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
Sinica} \textbf{33}, 104--107 (2018).}

\leavevmode\hypertarget{ref-Muller2014Mers}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
neutralizing antibodies in camels, Eastern Africa, 1983--1997.
\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

\leavevmode\hypertarget{ref-Xu2004Epidemiologic}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
1030--1037 (2004).}

\leavevmode\hypertarget{ref-Kasso2013Ecological}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
\textbf{2013}, e187415 (2013).}

\leavevmode\hypertarget{ref-Voigt2016Bats}{}%
\CSLLeftMargin{39. }
\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
a Changing World}. (Springer International Publishing, 2016).
doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}

\leavevmode\hypertarget{ref-Williams-Guillen2008Bats}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
\textbf{320}, 70--70 (2008).}

\leavevmode\hypertarget{ref-Gonsalves2013Mosquito}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
\emph{PLOS ONE} \textbf{8}, e77183 (2013).}

\leavevmode\hypertarget{ref-Gonsalves2013Mosquitoes}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
Research} \textbf{40}, 10--24 (2013).}

\leavevmode\hypertarget{ref-Mello2011Missing}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
\emph{PLOS ONE} \textbf{6}, e17395 (2011).}

\leavevmode\hypertarget{ref-Caraballo2022Crossspecies}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
Coronaviruses in the Americas: Contrasting Patterns between
Alphacoronavirus and Betacoronavirus. \emph{Microbiology Spectrum}
\textbf{0}, e01411--22 (2022).}

\leavevmode\hypertarget{ref-Wu2021Comprehensive}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
sarbecoviruses across China for the origin tracing of SARS-CoV and
SARS-CoV-2}. (2021)
doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}

\leavevmode\hypertarget{ref-Carlson2022Climate}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Alves2018Geographic}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
relationship between large-scale environmental determinants and bat
species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
(2018).}

\leavevmode\hypertarget{ref-Treitler2016Effect}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
effect of local land use and loss of forests on bats and nocturnal
insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}

\leavevmode\hypertarget{ref-Mendenhall2014Predictinga}{}%
\CSLLeftMargin{49. }
\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
E. A. \& Daily, G. C. Predicting biodiversity change and averting
collapse in agricultural landscapes. \emph{Nature} \textbf{509},
213--217 (2014).}

\leavevmode\hypertarget{ref-Amman2011Investigating}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-OHHLEP2022One}{}%
\CSLLeftMargin{51. }
\CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
sustainable and healthy future. \emph{PLOS Pathogens} \textbf{18},
e1010537 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}
